### 2018-05-27 evaluation.tex
Final revision from Prof. Xue

In our evaluation, we use the tiling factors
computed as in Section~\ref{subsec:example}:
$M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
The parallelism is controlled by two parameters,
the number of active clusters $NC$
and the number of active threads per cluster $NT_C$.
The total number of threads is $NT = NC \cdot NT_C$.
For the Phytium 2000+ processor, we have
$NC \in [1, 16]$ and $NT_C \in [1, 4]$. 
Generally, DGEMM performance is presented in $flops$.
As $NT$ varies in our evaluation,
we use another metric derived from $flops$,
the average thread efficiency,
to describe our performance results consistently.
The average thread efficiency is a normalized value computed as
$E_{avg} = flops / (NT \cdot \widehat{flops})$,
where $\widehat{flops}$ stands for the theoretical performance
peak of a single core.
%% \begin{equation*}
%%   E = flops / (NT \cdot \widehat{flops})
%% \end{equation*}

We show how cache sharing can hurt GEMM performance
(Section~\ref{subsec:drawback}),
demonstrate how SCP can reduce
the penalty caused by cache sharing 
(Section~\ref{subsec:benefit}), conduct a 
quantitative analysis of the cache miss rates
(Section~\ref{subsec:analysis}), and
discuss the privatization of shared matrix $B_2$
(Section~\ref{subsec:privb}).
%% Three versions of DGEMM are evaluated:
%% (1) Base, the baseline version,
%% (2) SCP,  the SCP version, 
%% and (3) SCP-P, the SCP version plus privatization of shared matrix $B_2$.

@@

As the L2 cache is shared by cores inside a cluster,
the impact of cache sharing can be demonstrated
by comparing the DGEMM performance results achieved
with the same $NT$
but different $NT_C$'s.
Figure~\ref{fig:drawback} shows the results.
Each curve represents a parallelism configuration.
Different subfigures are configured with different $NT$'s,
and different curves in the same subfigure are
configured with different $NT_C$'s.
As there is only one configuration for $NT=64$ ($NC=16$ and $NT_C=4$)
and no comparison is available,
the results for $NT=64$ and $NT=32$ are combined in Figure~\ref{fig:drawback-32}.

@@

We can see that for all $NT$'s, $NT_C=1$ achieves the 
best performance, followed by 
$NT_C=2$, and $NT_C=4$ comes at the lowest efficiency.
Table~\ref{tab:drawback} summarizes the average value
of $E_{avg}$ across all the matrix sizes.
Vertically, the results in the same column show how $E_{avg}$ varies with $NT_C$.
For all $NT$'s, $NT_C=1$ outperforms $NT_C=4$ by a large margin.
The performance gap is $7.4\%$ when $NT=4$ and roughly $10\%$ when $NT>4$.
Horizontally,  the results in the same row reflect the scalability of
DGEMM with certain $NT_C$'s.
With $NT_C=1$, the performance scales linearly with the number of threads.
With $NT_C=2$, $E_{avg}$ suffers a $1.86\%$ loss when $NT$ grows from 4 to 32.
The situation is much worse for $NT_C=4$, in which
$E_{avg}$ drops by $14.19\%$ from
$NT=4$ to $NT=64$.

This shows clearly that cache sharing has a considerable
impact on DGEMM performance.

@@

Under all the parallelism configurations, 
SCP performs better than Base
consistently over all the matrix sizes evaluated.
Table~\ref{tab:win} summarizes the performance improvement of SCP over Base.
%% Table~\ref{tab:benefit} summarizes the average value
%% of $E_{avg}$ in SCP evaluation.
%% There are two values in each cell, separated by a slash.
%% $E_{avg}$ stands on the left,
%% and performance win over conventional packing
%% in the corresponding parallelism configuration
%% is shown on the right for convenience.
With $NT_C=2$, SCP performs slightly better than Base
by $1.7\%$ -- $1.8\%$.
The performance gap becomes larger with $NT_C=4$,
ranging from $2.75\%$ to $6.91\%$.
The largest performance gain, $6.91\%$, which is a considerable improvement,
is observed under the maximum parallelism $NT=64$.
%% Looking at Table~\ref{tab:benefit} vertically,
%% we can see that SCP still cannot fully eliminate the impact
%% of cache sharing.
%% But it do relieve the situation to a certain degree.

@@

To understand why SCP outperforms Base and
why SCP is effective in reducing
inter-thread conflicts, 
we use the PAPI profiling tool~\cite{papi} to
analyze the cache behavior of GEMM. A total of 
four cache-related hardware events are counted,
which are listed and briefly described in Table~\ref{tab:events}.
Based on these hardware events, the
miss rates of L1 and L2 caches can be estimated  as 
$\frac{L1M}{L1M+L1H}$ and $\frac{L2M}{L2M+L2H}$, respectively.
The hardware events are measured in a per-thread manner
and all the results are the average measurements
over all the threads.

@@

Figure~\ref{fig:papi} shows the miss rates of L1 and L2 data caches.  We only give the results with $NT_C=4$
as SCP achieves the largest improvement with full cache sharing.
Table~\ref{tab:papi} summarizes the average reduction of the
cache miss rates across the matrix sizes used.
Under all the parallelism configurations,
both L1 and L2 cache miss rates are reduced under
SCP, 
by roughly $13\%$ for L1 and nearly $10\%$ for L2.

For the Phytium 2000+ processor whose L1 caches are private,
only L2 caches are specially handled by SCP.
The miss rates for both L1 and L2 caches
are reduced effectively by SCP.
While a reduction on the L2 miss rate is expected,
why is the L1 miss rate also reduced? In this processor,
every L2 cache is inclusive.
The conflict misses for the L2 cache not only cause
the data to be evicted from L2,
but also invalidate the evicted data's copy in all upper level L1 caches,
resulting in more L1 cache misses.
In general, conflict misses on an inclusive cache can affect
the performance of all upper level caches. As a 
result, improving the L2 miss rate also improves the
L1 miss rate.

Our results show that SCP can effectively
reduce the conflict misses in shared L2 caches.
In addition, the performance benefit thus obtained
also propagates to the upper level caches, too.

@@

\subsection{Privatization of Shared Matrix}\label{subsec:privb}
In the previous evaluations, SCP only set-partitions the shared cache
for $A_2$ matrices, which are private to threads.
The shared $B_2$ matrix is still packed
in the conventional way-partitioning style.
As mentioned in Section~\ref{subsec:example},
set-partitioning can also be applied to $B_2$ after $B_2$ is privatized.
In this section, we will discuss and evaluate this privatization alternative.

@@

Privatization incurs more packing overhead because
redundant packing of $B_2$ is enforced.
Figure~\ref{fig:packb} shows three packing strategies for $B_2$
in a parallelism configuration specified by
$NT_C=4$ and $NC=4$.
The whole packing workload of $B_2$ is divided into 16 tasks
and the packing workload of thread $T_1$ is masked by z-curves. 
%% and the task assignment is annoted in Figure~\ref{fig:packb}.
In the conventional packing approach, each thread takes a single task.
In full privatization, all tasks are replicated to all threads.
As a result, the packing overhead grows in proportion to $NT$
in a full privatization strategy, which is unacceptable
if GEMM is highly parallelized, e.g., when $NT=64$.
Partial privatization in Figure.~\ref{fig:packb.partial}
offers an alternative to full privatization.
In partial privatization, privatization only occurs inside
a cluster so that the tasks for one cluster need not to be
replicated to other clusters.
Consequently,  the extra packing overhead of the partial privatization
strategy is bounded by $NT_C$.
For example, on Phytium 2000+, overhead of packing $B_2$
in partial privatization is limited to 4 times
of that in conventional packing.

Figure~\ref{fig:privb} shows the results of the partial
privatization strategy with $NT_C=4$ and $NC=4$.
The results under other parallelism configurations are similar.
For comparison purposes,
SCP with and without privatization are both presented
(denoted as SCP-P and SCP in Figure~\ref{fig:privb}, respectively).
Figure~\ref{fig:privb.papi} shows the cache miss rates.
Privatization achieves slightly lower miss rates for
both L1 and L2 caches.
Figure~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower cache miss rates,
privatization suffers a performance loss compared to SCP.
Why is there a performance drop when cache miss rates 
are reduced?
The reason lies in the extra packing overhead introduced by privatization.
Figure~\ref{fig:privb.breakdown} shows the occupancy
ratios of the overall execution times for various overheads.
Overheads in GEMM include
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
While SCP-P involves roughly the same overheads in
packing $A_2$ and synchronization as SCP,
it spends much more time in packing $B_2$.
Table~\ref{tab:breakdown} gives a breakdown of the
execution time, indicating that the overhead for packing $B_2$
is multiplied by a factor of $NT_C=4$,
and the total overhead increases from $2.35\%$ to $4.72\%$.

\begin{figure}
  \centering
  \subfigure[Cache miss rates]{
    \label{fig:privb.papi}
    \includegraphics[width=.31\textwidth]{figures/privb-papi}
  }
  \subfigure[Average thread efficiency]{
    \label{fig:privb.ate}
    \includegraphics[width=.31\textwidth]{figures/privb-ate}
  }
  \subfigure[Executing time breakdown]{
    \label{fig:privb.breakdown}
    \includegraphics[width=.65\textwidth]{figures/privb-breakdown}
  }
  \caption{Performance of partial privatization with $NT_C=4$ and $NC=4$}
  \label{fig:privb}
\end{figure}

\begin{table}
  \centering
  \caption{Executing time breakdown with $NT_C=4$ and $NC=4$}
  \label{tab:breakdown}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{lcccc}
    \toprule
     & Sync & PackA & PackB & GEBP\\
    \midrule
    SCP   & 1.01 & 0.53 & 0.80 & 97.65 \\
    SCP-P & 1.00 & 0.52 & 3.20 & 95.28\\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

In our evaluation, we use the tiling factors
computed as in Section~\ref{subsec:example}:
$M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$,
as this combination causes GEMM to deliver the best performance 
when SCP is not used.
The degree of parallelism is controlled by two parameters,
the number of active clusters $NC$
and the number of active threads per cluster $NT_C$.
The total number of threads is $NT = NC \times NT_C$.
For the Phytium 2000+ processor, we have
$NC \in [1, 16]$ and $NT_C \in [1, 4]$. 
Generally, the DGEMM performance is measured in $flops$.
As $NT$ varies in our evaluation,
we also  use another metric derived from $flops$,
the average thread efficiency,
to describe our performance results consistently.
The average thread efficiency is a normalized value computed as
$E_{avg} = flops / (NT \times cdot \widehat{flops})$,
where $\widehat{flops}$ stands for the theoretical performance
peak of a single core.
%% \begin{equation*}
%%   E = flops / (NT \cdot \widehat{flops})
%% \end{equation*}

We show how cache sharing can hurt GEMM performance
(Section~\ref{subsec:drawback}),
demonstrate how SCP can reduce
the penalty caused by cache sharing 
(Section~\ref{subsec:benefit}), conduct a 
quantitative analysis of the cache miss rates
(Section~\ref{subsec:analysis}), and
discuss the privatization of the shared matrix $B_2$
(Section~\ref{subsec:privb}).
%% Three versions of DGEMM are evaluated:
%% (1) Base, the baseline version,
%% (2) SCP,  the SCP version, 
%% and (3) SCP-P, the SCP version plus privatization of shared matrix $B_2$.

@@

In a cluster, the L2 cache is shared by its four cores.
The impact of cache sharing can be demonstrated
by comparing the DGEMM performance results achieved
with the same $NT$
but different $NT_C$'s.
Figure~\ref{fig:drawback} shows the results.
Each curve represents a parallelization configuration.
Different subfigures are configured with different $NT$'s,
and different curves in the same subfigure are
configured with different $NT_C$'s.
As there is only one configuration for $NT=64$,
($NC=16$ and $NT_C=4$),
no comparison is available. Thus,
the results for $NT=64$ and $NT=32$ are combined in Figure~\ref{fig:drawback-32}.

@@

For all $NT$'s, $NT_C=1$ achieves the 
best performance, followed by 
$NT_C=2$, and $NT_C=4$ comes at the lowest efficiency.
Table~\ref{tab:drawback} summarizes the average value
of $E_{avg}$ across all the matrix sizes.
Vertically, the results in the same column show how $E_{avg}$ varies with $NT_C$.
For all $NT$'s, $NT_C=1$ outperforms $NT_C=4$ by a large margin.
The performance gap is $7.4\%$ when $NT=4$ and roughly $10\%$ when $NT>4$.
Horizontally,  the results in the same row reflect the scalability of
DGEMM with certain $NT_C$'s.
With $NT_C=1$, the performance scales linearly with the number of threads.
With $NT_C=2$, $E_{avg}$ suffers a $1.86\%$ loss when $NT$ grows from 4 to 32.
The situation is much worse for $NT_C=4$, in which
$E_{avg}$ drops by $14.19\%$ from
$NT=4$ to $NT=64$.

This shows clearly that cache sharing has a considerable
impact on the DGEMM performance.

@@

Under all the parallelization configurations, 
SCP performs better than Base
consistently across all the matrix sizes evaluated.
Table~\ref{tab:win} summarizes the performance improvement of SCP over Base.
%% Table~\ref{tab:benefit} summarizes the average value
%% of $E_{avg}$ in SCP evaluation.
%% There are two values in each cell, separated by a slash.
%% $E_{avg}$ stands on the left,
%% and performance win over conventional packing
%% in the corresponding parallelism configuration
%% is shown on the right for convenience.
With $NT_C=2$, SCP performs slightly better than Base
by $1.7\%$ -- $1.8\%$.
The performance gap becomes larger with $NT_C=4$,
ranging from $2.75\%$ to $6.91\%$.
The largest performance gain, $6.91\%$, which is a considerable improvement,
is observed under the maximum parallelism with $NT=64$.
%% Looking at Table~\ref{tab:benefit} vertically,
%% we can see that SCP still cannot fully eliminate the impact
%% of cache sharing.
%% But it do relieve the situation to a certain degree.

@@

To understand why SCP outperforms Base and
why SCP is effective in reducing
inter-thread cache conflicts, 
we use the PAPI profiling tool~\cite{papi} to
analyze the cache behavior of GEMM. A total of 
four cache-related hardware events are counted,
which are listed and briefly described in Table~\ref{tab:events}.
Based on these hardware events, the
miss rates of L1 and L2 caches can be estimated  as 
$\frac{L1M}{L1M+L1H}$ and $\frac{L2M}{L2M+L2H}$, respectively.
In this analysis,
the hardware events are measured in a per-thread manner
and all the results are the average measurements
over all the threads.

@@

Figure~\ref{fig:papi} shows the miss rates of L1 and L2 data caches.
We only give the results with $NT_C=4$
as SCP achieves the largest improvement with full cache sharing.
Table~\ref{tab:papi} summarizes their
average reductions across the matrix sizes used.
Under all the parallelization configurations,
both L1 and L2 cache miss rates are reduced under SCP, 
by roughly $13\%$ for L1 and nearly $10\%$ for L2.

For the Phytium 2000+ processor whose L1 caches are private,
only L2 caches are specially handled by SCP.
However,
the miss rates for both L1 and L2 caches
are reduced effectively by SCP.
While a reduction on the L2 miss rate is expected,
why is the L1 miss rate also reduced? In this processor,
every L2 cache is inclusive.
The conflict misses for the L2 cache not only cause
the data to be evicted from L2,
but also invalidate the evicted data's copy in all its
upper level L1 caches,
resulting in more L1 cache misses.
In general, the conflict misses for
an inclusive cache can affect
the performance of all its upper level caches.
As a result, improving the miss rate of a shared L2 cache
also improves the miss rates of all its upper level L1 caches.

Our results show that SCP can effectively
reduce the conflict misses in shared L2 caches.
In addition, the performance benefit thus obtained
also propagates to their upper level caches, too.

@@

\subsection{Privatization of Shared Matrix}\label{subsec:privb}
In the previous evaluations, SCP only set-partitions the shared caches
for $A_2$ matrices, which are private to threads.
The shared $B_2$ matrix is still packed
in the conventional way-partitioning style.
As mentioned in Section~\ref{subsec:example},
set-partitioning can also be applied to $B_2$ after $B_2$ is privatized.
In this section, we discuss and evaluate this privatization alternative.

@@

Privatization incurs more packing overhead because
the packing of $B_2$ must be done redundantly.
Figure~\ref{fig:packb} shows three packing strategies for $B_2$
in a parallelization configuration specified by
$NT_C=4$ and $NC=4$.
In each case, the entire packing workload of $B_2$ is divided into 16 tasks.
So only the packing workload in one thread, $T_1$, 
which is highlighted by the z-curves (as shown), is illustrated.
%% and the task assignment is annoted in Figure~\ref{fig:packb}.
In the conventional packing approach
(Figure~\ref{fig:packb.conventional}),
each thread takes a single task.
In full privatization
(Figure~\ref{fig:packb.full}),
all tasks are replicated to all
the threads.
As a result, the packing overhead grows in proportion to $NT$, which is unacceptable
if GEMM is highly parallelized, e.g., when $NT=64$.
In partial privatization 
(Figure~\ref{fig:packb.partial}), which is 
an alternative to full privatization,
privatization only occurs inside
a cluster so that the tasks for one cluster need not to be
replicated in the other clusters.
In this case, the extra packing overhead
is bounded by $NT_C$.
For example, on Phytium 2000+, the
overhead of packing $B_2$
in partial privatization is limited to 4 times
of that in the conventional packing approach.

Figure~\ref{fig:privb} shows the results for partial
privatization with $NT_C=4$ and $NC=4$.
The results for the other 
configurations are similar.
For comparison purposes,
SCP with and without privatization,
denoted as SCP-P and SCP, respectively,
are given in Figure~\ref{fig:privb}.
Figure~\ref{fig:privb.papi} shows the L1 and L2
cache miss rates, which are 
slightly lower with privatization (SCP-P) than without
(SCP). Figure~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower L1 and L2 cache miss rates,
privatization (SCP-P) suffers a performance loss
(relative to SCP) due to 
the extra packing overhead introduced by privatization.
Figure~\ref{fig:privb.breakdown} compares the
three types of major overheads in percentage terms 
over the total execution time, including
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
While behaving similarly as SCP
in (1) and (3), SCP-P is worse than SCP in (2).
To understand this further,
Table~\ref{tab:breakdown} gives a breakdown of each's
execution time in terms of (1) -- (3) and the
computation time of GEBP. As
the overhead for packing $B_2$
is multiplied by a factor of $NT_C=4$,
the total overhead has thus increased
from $2.35\%$ (under SCP) to $4.72\%$ (under SCP-P).

\begin{figure}
  \centering
  \subfigure[Cache miss rates]{
    \label{fig:privb.papi}
    \includegraphics[width=.31\textwidth]{figures/privb-papi}
  }
  \subfigure[Average thread efficiency]{
    \label{fig:privb.ate}
    \includegraphics[width=.31\textwidth]{figures/privb-ate}
  }
  \subfigure[Execution time breakdown]{
    \label{fig:privb.breakdown}
    \includegraphics[width=.65\textwidth]{figures/privb-breakdown}
  }
  \caption{Performance of partial privatization with $NT_C=4$ and $NC=4$}
  \label{fig:privb}
\end{figure}

\begin{table}
  \centering
  \caption{Execution time breakdown with $NT_C=4$ and $NC=4$}
  \label{tab:breakdown}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{lcccc}
    \toprule
     & Sync & PackA & PackB & GEBP\\
    \midrule
    SCP   & 1.01 & 0.53 & 0.80 & 97.65 \\
    SCP-P & 1.00 & 0.52 & 3.20 & 95.28\\
    \bottomrule
  \end{tabular}
\end{table}

### 2018-05-27 intro.tex
Final revision from Prof. Xue

\section{Introduction}\label{sec:intro}
Dense linear algebra libraries are the most fundamental software in
scientific and engineering computing domains.
Basic Linear Algebra Subprograms (BLAS)~\cite{blas}  defines a collection
of APIs which act as standard building blocks for dense matrix operations.
BLAS routines are divided into 3 levels,
level-1 for vector-vector operations,
level-2 for matrix-vector operations,
and level-3 for matrix-matrix operations.
Processor vendors often provide BLAS implementations
that are highly optimized for their processors,
such as Intel MKL, AMD ACML and NVIDIA cuBLAS.
The HPC community has also contributed several high-quality
BLAS implementations, e.g., ATLAS~\cite{atlas},
GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas},
and BLIS~\cite{blis,blisport}.

@@

%% GEMM performs a matrix-multiply-accumulate operation.
An optimized GEMM implementation consists of two components,
(1) a highly optimized kernel routine to accomplish its
computation, and
(2) an overall strategy to partition the workload into small tasks
and schedule these tasks to be executed effectively on the target processor.
The kernel routine is a serial program performing matrix-multiply-accumulate
on matrices from a single task.
The overall strategy partitions the workload by tiling 
the underlying loop nest
and determines a task schedule by choosing a specific traversal order
in the loop nest's iteration space.
In a multi-threaded context, different tasks (loop tiles) are assigned to
different threads to parallelize the whole operation.
The main objective of GEMM optimization is to
maximize the floating-point
operation throughput, measured by FLoating-point Operations Per Second ($flops$).
For the serial kernel routine, the optimization applies
instruction scheduling to improve instruction throughput.
For the workload partitioning, the optimization 
reorganizes memory accesses to reduce long memory latencies for
the kernel.
In this paper, we focus on optimizing the memory accesses for GEMM.

Two techniques, software prefetching and data packing,
have been widely used in GEMM implementations to speed up memory accesses.
Software prefetch instructions are utilized
to load data into cache before they are used, in 
order to ensure that memory accesses will not incur
long latencies and floating-point instructions
can execute at peak throughput.
To better exploit the capability of on-chip caches,
matrices are blocked and then packed into continuous memory buffers
that can fit into the caches under consideration
before the kernel routine is called.
Because GEMM is a computational intensive operation whose
arithmetic complexity is $O(n^3)$ and memory complexity is $O(n^2)$,
the overhead caused by data packing is negligible when
the input matrices are large~\cite{gotogemm}.

However, software prefetching and data packing cannot be effectively
applied to architectures with non-LRU shared caches.
In traditional high-performance processors, e.g., the Intel Xeon series,
both L1 and L2 caches are private to processor cores
and LRU replacement policy is adopted.
Without considering the last level cache (LLC),
the threads running on different cores prefetch their data into
different caches, and private data of different threads
cannot cause conflict cache misses.
This is no longer the case if caches are shared by several processor cores.
In this case, a cache line prefetched by one thread
may be evicted before its lifetime ends because another thread prefetches
another cache line into the same cache set.
A simple-minded solution would be to reduce 
the size of a packed matrix
so that the data used by several threads can simultaneously reside in a shared cache.
Due to the nature of set-associative caches,
this solution cannot completely eliminate inter-thread cache conflicts,
especially if the shared cache has a non-LRU replacement policy.
Our evaluation shows that inter-thread cache conflicts
in non-LRU shared caches
can heavily hurt GEMM performance,
even with packed matrices reduced to fit into the cache.

Following the trend in recent decades, modern architecture design
is introducing more and more cores on a single processor chip.
To reduce cost of on-chip cache memories and coherence networks,
shared caches may become common in future many-core processors.
In addition, non-LRU replacement policies, for example, pseudo-random,
may also be used to further reduce the cache design complexity.
Indeed, some high-performance processors based on the
ARM technology, e.g., the Phytium processor series~\cite{phytium},
%% I'm sure they have shared L2 caches,
%% but have not found any technical specification about the replacement policy
%% XGene from AppliedMicro~\cite{xgene},
%% ThunderX from Cavium~\cite{thunderx}
have already adopted this design. 
As a result, developing a better solution to reduce
the inter-thread cache conflicts 
for non-LRU shared caches is important.

In this paper, we present a Shared Cache Partitioning (SCP) method
to reduce inter-thread cache conflicts on architectures with non-LRU share caches.
The key idea is to partition a
share cache into physically disjoint sets
and assign different sets to different threads. This can be achieved by
exploiting the memory mapping mechanism of set-associative caches.

To the best of our knowledge, SCP represents
the first work addressing
the inter-thread cache conflicts problem on
architectures with non-LRU share caches.

The main contributions of this paper are as follows:
\begin{itemize}
\item We present a quantitative analysis of the negative effect of inter-thread cache
  conflicts on GEMM performance.
\item We propose SCP, a method for solving the inter-thread cache conflicts
  problem on architectures with non-LRU shared caches.
\item We have implemented SCP in the OpenBLAS library and evaluated it on Phytium 2000+,
  an emerging high-performance processor based on ARM's AArch64 architecture.
  Phytium 2000+ has 64 cores, private LRU L1 caches
  and pseudo-random shared L2 caches.
  Our evaluation shows that SCP can improve the GEMM performance consistently
  under various parallelism configurations, by $2.75\%$--$6.91\%$.
\end{itemize}

The rest of the paper is organized as follows.
Section~\ref{sec:background} gives some background on GEMM and discusses
the cache conflicts incurred.
Section~\ref{sec:scp} introduces the SCP method.
Section~\ref{sec:evaluation} presents and
analyzes our experimental results.
Section~\ref{sec:related} reviews the related work.
Finally, Section~\ref{sec:conclusion} concludes.

>>>>

\section{Introduction}\label{sec:intro}
Dense linear algebra libraries are the most fundamental software in
scientific and engineering computing domains.
Basic Linear Algebra Subprograms (BLAS)~\cite{blas}  
define a collection
of APIs which act as standard building blocks for dense matrix operations.
BLAS routines are divided into three levels,
level-1 for vector-vector operations,
level-2 for matrix-vector operations,
and level-3 for matrix-matrix operations.
Processor vendors often provide BLAS implementations
that are highly optimized for their processors,
such as Intel MKL, AMD ACML, and NVIDIA cuBLAS.
The HPC community has also contributed several high-quality
BLAS implementations, e.g., ATLAS~\cite{atlas},
GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas},
and BLIS~\cite{blis,blisport}.

@@

%% GEMM performs a matrix-multiply-accumulate operation.
An optimized GEMM implementation consists of two components,
(1) a highly optimized kernel routine to accomplish its
computation, and
(2) an overall strategy to partition the workload into small tasks
and schedule these tasks to be executed effectively on the target processor.
The kernel routine is a serial program performing matrix-multiply-accumulate
on matrices from a single task.
The overall strategy partitions the workload by tiling 
the underlying loop nest
and determines a task schedule by choosing a specific traversal order
in the loop nest's iteration space.
In a multi-threaded context, different tasks (loop tiles) are assigned to
different threads to harness parallelism.
The main objective of the GEMM optimization is to
maximize the floating-point
operation throughput, measured by FLoating-point Operations Per Second ($flops$).
For the serial kernel routine, the optimization applies
instruction scheduling to improve instruction throughput.
For the workload partitioning, the optimization 
reorganizes memory accesses to reduce long memory latencies for
the kernel.
In this paper, we focus on optimizing the memory accesses for GEMM.

Two techniques, software prefetching and data packing,
have been widely used in GEMM implementations to speed up memory accesses.
Software prefetch instructions are utilized
to load data into cache before they are used, in 
order to ensure that memory accesses will not incur
long latencies and floating-point instructions
can execute at peak throughput.
To better exploit the capability of on-chip caches,
matrices are blocked and then packed into continuous memory buffers
that can fit into the caches under consideration
before the kernel routine is called.
Because GEMM is a computational intensive operation 
with an 
arithmetic complexity of $O(n^3)$ and a memory complexity of $O(n^2)$,
the overhead caused by data packing is negligible for
large
input matrices~\cite{gotogemm}.

However, software prefetching and data packing cannot be effectively
applied to architectures with shared non-LRU caches.
In traditional high-performance processors, e.g., the Intel Xeon series,
both L1 and L2 caches are private to processor cores
and an LRU replacement policy is adopted.
Without considering the last level cache (LLC),
the threads running on different cores prefetch their data into
different caches, and private data of different threads
cannot cause conflict cache misses.
This is no longer the case if a cache is
shared by several processor cores.
In this case, a cache line prefetched by one thread
may be evicted before its lifetime ends because another thread prefetches
another cache line into the same cache set.
A simple-minded solution would be to reduce 
the size of a packed matrix
so that the data used by several threads can simultaneously reside in a shared cache.
Due to set-associativity,
this solution is usually ineffective in reducing
inter-thread cache conflicts,
especially if the shared cache has a non-LRU replacement policy.
Indeed,
our evaluation shows that inter-thread cache conflicts
in shared non-LRU caches
can heavily hurt GEMM performance,
even if packed matrices are reduced (in size)
to fit into the shared cache.

Following the trend in recent decades, modern architecture design
is introducing more and more cores on a single processor chip.
To reduce the cost of on-chip cache memories and coherence networks,
shared caches may become common in future many-core processors.
In addition, non-LRU replacement policies, for example, pseudo-random,
may also be used to further reduce the cache design complexity.
Indeed, some high-performance processors based on the
ARM technology, e.g., the Phytium processor series~\cite{phytium},
%% I'm sure they have shared L2 caches,
%% but have not found any technical specification about the replacement policy
%% XGene from AppliedMicro~\cite{xgene},
%% ThunderX from Cavium~\cite{thunderx}
have already adopted this design. 
As a result, developing a better solution to reduce
the inter-thread cache conflicts 
for shared non-LRU caches is important.

In this paper, we present a Shared Cache Partitioning (SCP) method
to reduce inter-thread cache conflicts on architectures with shared non-LRU caches.
The key insight is to partition a
shared cache into physically disjoint sets
and assign different sets to different threads. This can be achieved by
exploiting the memory mapping mechanism in a 
set-associative cache.

To the best of our knowledge, SCP represents
the first work addressing
the inter-thread cache conflicts problem on
architectures with shared non-LRU caches.

The main contributions of this paper are as follows:
\begin{itemize}
\item We present a quantitative analysis of the negative effect of inter-thread cache
  conflicts on the GEMM performance.
\item We propose SCP, a method for solving the inter-thread cache conflicts
  problem on architectures with shared non-LRU caches.
\item We have implemented SCP in the OpenBLAS library and evaluated it on Phytium 2000+,
  an emerging high-performance processor based on ARM's AArch64 architecture.
  Phytium 2000+ has 64 cores, private LRU L1 caches
  and shared L2 caches with a pseudo-random
  replacement policy.
  Our evaluation shows that SCP can improve the GEMM performance consistently
  under various parallelization configurations, by $2.75\%$--$6.91\%$.
\end{itemize}

The rest of the paper is organized as follows.
Section~\ref{sec:background} gives some background on GEMM and discusses
the cache conflicts incurred.
Section~\ref{sec:scp} introduces our SCP method.
Section~\ref{sec:evaluation} presents and
analyzes our experimental results.
Section~\ref{sec:related} reviews the related work.
Finally, Section~\ref{sec:conclusion} concludes.
